---
title: "Thrive Market Analysis, Part II"
author: "Mayra Pérez"
date: 2022-06-20
categories: ["R"]
tags: ["Data origins", "limitations", "EDA"]
---

## The Data Set: Origins

Data is everywhere, yet there comes a time when the information we need isn't readily available. What does one do when the details we need are out of reach? *One scrapes the web.* For inexperienced hackers, this translates into hours of research and implementation. Do expect loss of emotional regulation and bloodshot eyes.

The following data set was sourced from Thrive Market, an e-commerce retailer focused on natural and organic foods, using web scraping methods. It includes information on approximately 322 condiments and sauces. There are about 16 features for each product from Brand to ingredients. Below are descriptions of each feature.

- **id**: *(character)* unique product identifier presented as a numerical value
- **name**: *(character)* product name
- **brand**: *(character)* product manufacturer
- **description**:*(character)* detailed product description
- **price**: *(double class)* product price, which may be less than the Manufacturer's Suggested Retail Price (MSRP)
- **size**: *(character)* product size in terms of weight and/or packaging
- **value_awards**: *(character)* metadata on awards conferred, such as Certified Organic
- **values_diet**: *(character)* metadata on diet values, such as Vegan
- **values_social**: *(character)* metadata on social values, such as AAPI-owned
- **values_health**: *(character)* metadata on health values, such as Gluten-Free
- **sales_rank**: *(double class)* sales rank based on how well the product is selling relative to other products in the same category
- **ratings_rank**: *(double class)* product rank based on how well the product is valued or rated relative to other products in the same category
- **rating_count**: *(double class)* number of ratings
- **rating_avg**: *(double class)* average rating score
- **review_score**: *(double class)* average review score
- **review_count**: *(double class)* number of reviews
- **allergens**: *(character)* list of allergens
- **ingredients**: *(character)* list of product ingredients

## The Data Set: Limitations

### Size

The first limitation of the Thrive Market data set is its small size. At 322 observations, this represents *all* available data on products in the condiments and sauces category. Nevertheless, there are some important considerations as I continue Exploratory Data Analysis (EDA) and, eventually, modeling.

### Bias

The second limitation is related to the first. In the article *Reducing the Effect of Sample Bias for Small Data Sets with Double-Weighted Support Vector Transfer Regression*, the authors explain "small data sets are an extremely challenging problem in the machine learning (ML) realm, and in specific, in regression scenarios, as the lack of relevant data can lead to ML models that have large bias" (Luo and Paal, 2020). Google Chief Decision Scientist, Cassie Kozyrkov, helps conceptualize this frequently talked about, but complex term as "systematic lopsidedness" which means the results can be misleading (Kozyrkov, 2020). 

For my particular data set, what follows are the most relevant types of biases and ways to address them (Suresh and Guttag, 2019):

#### *Representation Bias*

Representation bias is the result of a data set that does not evenly represent all groups. Some groups may be over- or under-represented. Representation bias is the biggest concern for this project.

The Thrive Market data set certainly represents all products within a specific category at the e-commerce site. However, the data set is not balanced. For instance, some brands are over-represented while others are under-represented. As I transition to model development, it may be beneficial to consider the effect of Brand over- or under-representation on overall popularity of a product. A product may be popular because it's widely available on the site and, therefore, successful due to higher brand awareness.

One method to mitigate representation bias is sample weights. In other words, assigning a value that is inversely proportional to the frequencies of each group.

#### *Aggregation Bias*

Aggregation bias sometimes relates to representation bias. It is the result of assuming features have the same effect or implications on all groups.

Given the possibility of representation bias in the Thrive Market data set, it will be critical to be mindful of aggregation bias. The challenge will be to identify how the data are different and create separate models according to group classification. This may addressed at the beginning by creating a representative sample or transforming unbalanced predictors.

#### *Evaluation Bias*

Evaluation bias is the result of methods that are not applicable to a target population. For the purposes of this project, evaluation bias may less of a concern.

Thrive Market attracts health- and planet-conscious consumers. It will be important to highlight that the results of this project may not be applicable to products aimed at another type of consumer. For instance, in answering the research question (which features are predictive of customer satisfaction?), the insights may not be relevant to products sold at other e-commerce retailers, such as Amazon or Wal-Mart.

### Quality

The third limitation is the untidiness of the "raw" Thrive Market data set. However, this is a common issue in data science. To date, I have devoted ~85% of my time on this project preparing the data set for engineering and modeling. For example, the various categories of customer values were indistinguishable. It was difficult to differentiate between awards, diet, and social values. Categories that appeared in one column, such as Gluten-Free, for one product, appeared in another column for other products. Consequently, I collapsed all values into a single values category as these appear on the e-commerce site under a single category.

## Exploratory Data Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
thrive_products <- read_excel("~/Desktop/Willamette MBA/8_Summer2022/DS Capstone/Final Project/1. EDA/thrive-2.xlsx")
library(tidyverse)
thrive_products <- thrive_products %>%
  select(-...1)

thrive_products$price <- as.numeric(gsub("\\$", "", thrive_products$price)) ## Remove dollar sign and transform to numeric
```

#### There are 58 brands in the data set.

```{r, include = FALSE}
all_brands <- thrive_products %>%
  count(brand)
```

```{r}
all_brands
```

#### Some brands are over-represented.

```{r, include = FALSE}
brands_viz <- thrive_products %>%
  count(brand) %>%
  top_n(5)
```

```{r, include = FALSE}
top_brands <- ggplot(brands_viz, mapping = aes(x = reorder(brand, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top brands by products available", subtitle = "About 42% of products are sold by 6 companies") +
  xlab("Brand") +
  ylab("Count") +
  theme_bw()
```

```{r}
top_brands
```

#### Price is positively skewed.

```{r add missing prices, include = FALSE}
thrive_products$price[thrive_products$id == "15197"] <- 19.99
thrive_products$price[thrive_products$id == "23848"] <- 41.98
thrive_products$price[thrive_products$id == "23847"] <- 47.97
thrive_products$price[thrive_products$id == "23849"] <- 19.98
```

```{r, include = FALSE}
price <- ggplot(thrive_products, aes(price)) +
  geom_histogram()
```

```{r}
price
```

```{r, include = FALSE}
thrive_products <- thrive_products %>% 
  mutate(lprice = log(price))
```

#### A log transformation somewhat normalizes the data.

```{r, include = FALSE}
l_price <- ggplot(thrive_products, aes(lprice)) +
  geom_histogram()
```

```{r}
l_price
```

```{r, include = FALSE}
## Create a subset of values
thrive_products$values <- paste(thrive_products$values_awards,
                            thrive_products$values_diet,
                            thrive_products$values_health,
                            thrive_products$values_social,
                            sep = ", ")

thrive_values <- thrive_products %>%
  select(id, name, brand, values)

thrive_values <- separate_rows(thrive_values, values, sep = ",") ## separate values
thrive_values$values <- str_trim(thrive_values$values, side = c("both", "left", "right")) ## remove white space from all sides
```

```{r, include = FALSE}
## Top 5 values
values_viz <- thrive_values %>%
  count(values) %>%
  top_n(5)

values <- ggplot(values_viz, mapping = aes(x = reorder(values, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top diet, social, and health values", subtitle = "About 98% of products have a Non-GMO pledge") +
  xlab("Value") +
  ylab("Count") +
  theme_bw()
```

#### Thrive Market delivers on healthy values. There's something for everyone.

```{r}
values
```

```{r, include = FALSE}
thrive_products$sales_rank <- as.numeric(thrive_products$sales_rank) ## Transform to factor? Should this be continuous or discrete?

thrive_products$sales_rank <- as.numeric(thrive_products$sales_rank) ## Transform to factor? Should this be continuous or discrete?

thrive_products$ratings_rank <- as.numeric(thrive_products$ratings_rank) ## Transform to factor? Should this be continuous or discrete?

thrive_products$rating_count <- as.numeric(thrive_products$rating_count) ## Transform to numeric; count should be continuous

thrive_products$rating_avg <- as.numeric(thrive_products$rating_avg) ## Transform to numeric; average should be continuous

thrive_products$review_score <- as.numeric(thrive_products$review_score) ## Transform to numeric; score should be continuous

thrive_products$review_count <- as.numeric(thrive_products$review_count) ## Transform to numeric; count should be continuous
```

```{r, include = FALSE}
average_rating <- ggplot(thrive_products, aes(rating_avg)) +
geom_histogram() +
  xlab("Average Rating") +
  ylab("Count") +
  theme_bw()
```

#### The vast majority of products are rated between 3 and 5.

```{r}
average_rating
```

#### Sales and ratings ranks are stored as numeric values, but perhaps they should be stored as factors. These labels are used to categorize products from low to high performers.

```{r, include = FALSE}
sales_rank <- ggplot(thrive_products, aes(sales_rank)) +
geom_histogram() +
  xlab("Product Sales Rank") +
  ylab("Count") +
  theme_bw()
```

```{r, include = FALSE}
ratings_rank <- ggplot(thrive_products, aes(ratings_rank)) +
geom_histogram() +
  xlab("Product Ratings Rank") +
  ylab("Count") +
  theme_bw()
```

```{r}
sales_rank
```

```{r}
ratings_rank
```

#### That's all folks! Stay tuned for insights on product descriptions, allergens, ingredients and other untidy data.


## References

Kozyrkov, C., 2020. What is AI bias?. [online] Medium. Available at: <https://towardsdatascience.com/what-is-ai-bias-6606a3bcb814> [Accessed 15 June 2022].

Luo, H. and Paal, S., 2020. Reducing the effect of sample bias for small data sets with double-weighted support vector transfer regression. Computer-Aided Civil and Infrastructure Engineering, 36(3), pp.248-263.

Suresh, H. and Guttag, J.V., 2019. A framework for understanding unintended consequences of machine learning. arXiv preprint arXiv:1901.10002, 2, p.8.